---
title: "Get started with emr: Procedure"
author: "Hsiang-Ju, Chiu"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Intro: Pr}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage{UTF-8}{inputenc}
---

# Description
```
cat(utils::packageDescription("emr")$Description)
```
### Development version
```r
install.packages("devtools")
# Install development version from GitHub
devtools::install_github("DHLab-CGU/emr")
library(emr)
```
```{r setup, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(collapse = T, comment = "#>")
options(tibble.print_min = 4L, tibble.print_max = 4L)
devtools::install_github("DHLab-CGU/emr")
library(emr)
```
### ICD-CM code with two format

There are two formats of ICD-9 codes, decimal (with a decimal place separating the code) and short format.

ICD-9-PCS
```{r}
# ICD-9-PCS_Short
head(ICD9PrwithTwoFormat$Short)

# ICD-9-PCS_Decimal
head(ICD9PrwithTwoFormat$Decimal)
```
ICD-10-PCS只有short的格式
```{r}
# ICD-10-PCS_Short
head(prICD10$ICD)
```
### Sample file
An example of the data shows as below:  
This data is a simulated medical data set of 3 patients with 170 records.
```{r}
head(samplePrFile)
```
## I. Code standardization

**ICD-9-PCS code format transformation：Short <-> Decimal**

Standardize the ICD procedure codes into a uniform format before code classification.

Since the formats of ICD code used in the grouping tables could be different, the transformational type is according to the grouping method chosen by users.  

**Table 1** Format of code classification methods

|   |ICD format|
|--------|----|
|Clinical Classifications Software|short format|
|Procedure class|short format|

To distinguish the version of ICD procedure code (ICD9/ICD10) in code standardization.  
User can set the using date of ICD10.  

Reimbursement claims with a date required to use of ICD-10 codes (`icd10usingdate`)   
In America: 10/1/2015  
In Taiwan: 1/1/2016   

**A. Uniform decimal format**

```{r, message = FALSE, warning = FALSE}
decimal <- icdPrShortToDecimal(prDataFile = samplePrFile,
                               icdColName = ICD, 
                               dateColName = Date,
                               icd10usingDate = "2015/10/01")
head(decimal$ICD)
```
**B. Uniform short format**
Convert the diagnostic codes to the short format, and used for grouping to the others classification (`icdPrToCCS` and `icdPrToCCSLvl`). 
```{r, message = FALSE, warning = FALSE}
short <- icdPrDecimalToShort(prDataFile = samplePrFile,
                             icdColName = ICD, 
                             dateColName = Date,
                             icd10usingDate = "2015/10/01")
head(short$ICD)
```
**warning message** (More detail: [Get started with emr: Diagnosis](https://dhlab-cgu.github.io/emr/articles/Eng_Diagnosis.html#i--code-standardization))

Through code standardization functions generate data of procedure codes with potential error, to help a researcher identify the potential coding mistake that may affect clinical data analysis.

## II. Data integration
These functions collapse ICD codes into a smaller number of clinically meaningful categories that are more useful for presenting descriptive statistics than are individual ICD procedure codes.

The emr package supports two strategies to group EHR procedure codes, including CCS and procedure classes.  

### A. Clinical Classifications Software (CCS)
The CCS classification for ICD-9 and ICD-10 codes is a procedure categorization scheme that can employ in many types of projects analyzing data on procedures. 

**1) single-level**: `icdPrToCCS`

ICD-9-PCS and ICD-10-PCS code contains same 231 CCS categories.

```{r, message = FALSE, warning = FALSE}
## ICD to CCS category
CCS <- icdPrToCCS(prDataFile = samplePrFile,
                  idColName = ID,
                  icdColName = ICD,        
                  dateColName = Date,
                  icd10usingDate = "2015-10-01",
                  isDescription = FALSE)

head(CCS$groupedDT, 5)
```

**2) multi-level**: `icdPrToCCSLvl`

ICD-9-PCS has 3 levels, and ICD-10-PCS has two levels

```{r, message = FALSE, warning = FALSE}
## ICD to CCS multiple level 2 description
CCSLvl <- icdPrToCCSLvl(prDataFile = samplePrFile,
                       idColName = ID,
                       icdColName = ICD,        
                       dateColName = Date,
                       icd10usingDate = "2015-10-01",
                       CCSLevel = 2,
                       isDescription = TRUE)

head(CCSLvl$groupedDT, 5)
```

### B. Procedure Class 
Procedure Classes are part of the family of databases and software tools developed as part of the HCUP by AHRQ.     
The Procedure Classes assign all ICD procedure codes to one of four categories:  

* Minor Diagnostic - Non-operating room procedures that are diagnostic 
* Minor Therapeutic - Non-operating room procedures that are therapeutic
* Major Diagnostic - Operating room procedures that are diagnostic
* Major Therapeutic - Operating room procedures that are therapeutic
 
```{r, message = FALSE, warning = FALSE}
ProcedureClass <- icdPrToProcedureClass(prDataFile = samplePrFile,
                                        idColName = ID,
                                        icdColName = ICD,      
                                        dateColName = Date,
                                        icd10usingDate = "2015-10-01",
                                        isDescription = FALSE)
```
```r
head(ProcedureClass$groupedDT, 5)
#>    Short ID   ICD       Date PROCEDURE_CLASS_NAME
#> 1:  5681  B  5681 2008-01-14                    4
#> 2:  9774  A  9774 2009-01-11                    2
#> 3:  4499  B 44.99 2009-05-10                    4
#> 4:  0759  C 07.59 2009-01-21                    4
#> 5:  0205  B  0205 2008-07-06                    4
```

## Reference

### I. Code standardization

ICD-9-PCS code (2014): https://www.cms.gov/Medicare/Coding/ICD9ProviderDiagnosticCodes/codes.html

ICD-10-PCS code (2019):https://www.cms.gov/Medicare/Coding/ICD10/2019-ICD-10-PCS.html

https://www.findacode.com/search/search.php

https://www.cms.gov/Medicare/Quality-Initiatives-Patient-Assessment-Instruments/HospitalQualityInits/Downloads/HospitalAppendix_F.pdf

### II. Data integration

** Clinical Classifications Software (CCS)**

ICD-9-PCS CCS (2015): https://www.hcup-us.ahrq.gov/toolssoftware/ccs/Single_Level_CCS_2015.zip

https://www.hcup-us.ahrq.gov/toolssoftware/ccs/Multi_Level_CCS_2015.zip

ICD-10-PCS CCS (2019): https://www.hcup-us.ahrq.gov/toolssoftware/procedureicd10/procedure_icd10.jsp

**Procedure Class**

ICD-9-Procedure Class (2015): https://www.hcup-us.ahrq.gov/toolssoftware/procedure/pc2015.csv

ICD-10-Procedure Class (2019): https://www.hcup-us.ahrq.gov/toolssoftware/procedureicd10/procedure_icd10.jsp
